---
title: "Class 1"
author: "DABAJA"
date: "9/13/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Instal packages
```{r Packages}
#install.packages(c("timeDate","mvtnorm","quadprog","quantreg","rgl","robustbase","scatterplot3d","SparseM","tseries"))
#install.packages("/Users/dawidjarosz/Dropbox/Semester\ 11/Statistical\ Analysis\ of\ Financial\ Data/Class\ 1/Rsafd", repos = NULL, type="source")

library(timeDate)
library(mvtnorm)
library(quadprog)
library(rgl)
library(robustbase)
library(scatterplot3d)
library(SparseM)
library(tseries)
library(Rsafd)
```

Load the data set: Rdata rdb - file with databases he works with in the book. 
```{r}
lazyLoad(filebase = "/Users/dawidjarosz/Dropbox/Semester\ 11/Statistical\ Analysis\ of\ Financial\ Data/Class\ 1/Rsafd/data/Rdata", 
         envir = parent.frame())
```

```{r}
wn <- rnorm(1024)
```
Samuelson 
```{r}
DELTAT <- 1/252
SIG <- .2*sqrt(DELTAT)
MU <- .15*DELTAT
TIME <- (1:1024)/252
STOCK <- rep(0,1024) #repeats a zero for every entry 
```

Model
```{r}
for (i in 1:1024) {
  STOCK[i] <- exp(SIG*RW[i] + MU*TIME[i]) # this is exponential 
}
```

without a loop 

```{r}
STOCK <- exp(SIG*RW + MU*TIME)
```

Plot
```{r}
plot(TIME, STOCK, type="l")
```



```{r}
head(HOWAREYOU)
```

Building function, Black and Scholes - 

```{r}
S <- 100 #stock price 
K <- S 
R <- .4 #per year
SIG <- .2 #annualized vol 

TAU <-90/252  #time to maturity, trading days. 

d1 <- log(S/K) + TAU * (R+SIG^2/2)
d2 <- d1/(SIG*sqrt(TAU))
d2 <- d1 - SIG*sqrt(TAU)
```

Making a BS funciton, pricing a call. 

bs_call <- function() #locally 

fix(bs_call) #for any other script, use it a lot, this is how you'll have access to it later

# the function doesn't work 
```{r}
bs_call <- function(S=100, 
                    K=100,
                    R=.1,
                    SIG=2,
                    TAU=90/252){
  d1 <- log(S/K) + TAU * (R+SIG^2/2)
  d1 <- d1/(SIG*sqrt(TAU))
  d2 <- d1 - SIG*sqrt(TAU)

  c <- s*pnorm(d1) - K*exp(-R*TAU)*pnorm(d2)
  return(c)                  }
```

```{r}
#bs_call(S=109, K=109, R=0.2)
```

Precept
```{r Objects and matrices}
# Check/change working directory:
getwd()
#setwd("path...")
# can also be done using ctrl+shift+H


#### Creating R objects ####
# Creates a vector of length 16 containing the first 16 integers in increasing order:
x <- 1:16

# Same:
y <- seq(from=1, to=16, by=1)

# To know what the function seq does:
help(seq)

# If you type the name of the function or method without parentheses, 
# the system will return the R code of the function in question:
objects() # lists objects in the workspace
objects   # gives code for this function

dim(x) <- c(8,2)  # Organizes the 1-dim vector in a 8x2 matrix
                  # the function c is used for concatenation
class(x)  # all objects created by R have a class atribute, which can be retrieved
          # by using the function class

z <- x*x  # creates a new R object which is also a 8x2 matrix obtained by multiplying
          # entry by entry
# Z <- x%*%x    # usual matrix product ('Error: non-conformable arguments' means we 
              # can't multiply these two matrices due to dimension problems)
Z1 <- x%*%t(x)  # 8x8 matrix
Z2 <- t(x)%*%x  # 2x2 matrix
# t stands for transpose of the matrix
```


```{r}
rm(x) # removes objects; used to save memory 
rm(y,z)

dump(c("Z1", "Z2"), "mymatrices")   # dump produces text representations of objects
# dump needs 2 parameters: vectors of names of objects to be found in the workspace,
# and file name to dump the representations to. Outside of R!
```


## Random Generation and White Noise ##

```{r}
WN <- rnorm(1024)     # create vector of length 1024
help(rnorm)
WN2 <- rnorm(1024,5.0,4.0) #mean and sd modified 

plot(WN, type = "l")  # produce sequential plot of the entries of this vector

graphics.off()
par("mar")
par(mar=c(1,1,1,1))


par(mfrow = c(2,1))
# mfrow=c(nrows, ncols) to create a matrix of nrows x ncols plots that are filled
# in by row. mfcol=c(nrows, ncols) fills in the matrix by columns.
plot(WN,type="l")
plot(WN[1:64],type="l") # notice the slicing
```

#### More functions and for loops ####
```{r}
RW <- cumsum(WN) # cumsum is the discrete analogue to integration, just as diff is
                 # the discrete analogue to differentiation
plot(RW,type="l") # Sample of size 1024 from a Random Walk
```

# Building Models #

```{r Samualson's model}
# Create an object from Samualson's model for stocks:
DELTAT <- 1/252
SIG <- .2*sqrt(DELTAT)
MU <- .15*DELTAT
TIME <- (1:1024)/252  # expressed in years, there are 252 trading days
STOCK <- rep(0,1024)  # initial zero vector for the stock

for (I in 1:1024) 
  STOCK[I] <- exp(SIG*RW[I]+MU*TIME[I])

# To avoid the 'for' loop (and save time):
STOCK <- exp(SIG*RW+MU*TIME)

plot(RW, type="l")
title("Sample of size 1024 from a Random Walk")
plot(TIME, STOCK,type="l")
title("Corresponding Geometric Random Walk with Drift")
```

#### Importing Data ####
```{r}
# Installing packages (please download the Rsafd file first):
install.packages(c("timeDate","mvtnorm","quadprog","quantreg","rgl","robustbase","scatterplot3d","SparseM","tseries"))


# Fix libraries that are used very frequently:
#.First()
#fix(.First)
# library(Rsafd)
# cat("\n Welcome to my world\n\n")
```
```{r}
#attach(hills) # see item in INDEX file
#plot(climb,time)
#plot(dist,time)
```
#### Importing data from .rdb file ####
```{r}
setwd("/Users/dawidjarosz/Dropbox/Semester\ 11/Statistical\ Analysis\ of\ Financial\ Data/Rsafd/data")
readRDS("Rdata.rds")  # see metadata


lazyLoad(filebase = "/Users/dawidjarosz/Dropbox/Semester\ 11/Statistical\ Analysis\ of\ Financial\ Data/Rsafd/data/Rdata",
         envir = parent.frame())

#find(HOWAREYOU)
head(HOWAREYOU)
```

#### Getting data from the Web via Excel ####
Don't have the "toronto" file.

```{r}
#toronto <- read.csv("path...\\file.csv")

#class(toronto)
#dim(toronto)
#head(toronto) # notice the reversed order
#tail(toronto) 
#toronto.close <- toronto[,4]  # closing price is in the 4th column
#plot(toronto.close,type="l")
#tsx <-rev(toronto.close)  # reverses the order of the data
#plot(tsx,type="l") 
```

#### First Function ####

# e.g. Black-Scholes formula
```{r}
S <- 100  # price of the underlying asset
K <- S  # at-the-money option
R <- .1  # interest rate
SIG <- .2  #annualized volatility
TAU <- 90/252  # time to maturity (90 days) in years

d1 <- log(S/K) + TAU*(R+SIG^2/2)
d1 <- d1/(SIG*sqrt(TAU))
d2 <- d1 - SIG*sqrt(TAU)

C <- S*pnorm(d1) - K*exp(-R*TAU)*pnorm(d2)

```

# Building a function:

# 1) Create a .txt file that contains the function and then call
# it using 'source'.


\textit{ How to save it into a file that I can later use and upload?}

```{r}
# source('bs_call_test.txt')
```


# 2) Local function:
```{r}
bs_call <- function (TAU=90/252,K=100,S=100,R=.1,SIG=.2) 
{
  # Parameters:   TAU - time to maturity in years
  #               K - strike
  #               S - current value of the underlying
  #               R - yearly interest rate
  #               SIG - annualized volatility
  # Return: Black-Scholes call price
  
  d1 <- log(S/K) + TAU*(R+SIG^2/2)
  d1 <- d1/(SIG*sqrt(TAU))
  d2 <- d1 - SIG*sqrt(TAU)
  
  C <- S*pnorm(d1) - K*exp(-R*TAU)*pnorm(d2)
  return(C)
  
}
```


# 3) Global function:
```{r}
fix(bs_call)  # define a function by copying the local function above on the pop-up window



# Using the function:
bs_call() # use the function with default values
bs_call(TAU=60/252) # change one or more of the default values
```

