---
title: "Class 3"
author: "DABAJA"
date: "9/27/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(timeDate)
library(mvtnorm)
library(quadprog)
library(rgl)
library(robustbase)
library(scatterplot3d)
library(SparseM)
library(tseries)
library(Rsafd)
```


```{r}
lazyLoad(filebase = "/Users/dawidjarosz/Dropbox/Semester\ 11/Statistical\ Analysis\ of\ Financial\ Data/Rsafd/data/Rdata", 
         envir = parent.frame())
```

\section{Generalized Pareto Distribution}

GPD - In statistics, the generalized Pareto distribution (GPD) is a family of continuous probability distributions. It is often used to model the tails of another distribution. It is specified by three parameters: location $\mu$ , scale $\sigma$ , and shape  $\xi$ . Sometimes it is specified by only scale and shape and sometimes only by its shape parameter. Some references give the shape parameter as $\kappa = - \xi$ 

We plotting 3 one sided pareto distributions. To check how side parameters affect the... 

```{r}
x <- seq(from=.2, to=15, length=5000)
```

Ploting the shappes of various GPDs and the $\xi$ is being modifiad. 
```{r Graphing GPD}
plot(x, dpareto(x,xi=.2), type="l", ylab="GPD densities", ylim=c(-.05,1.1))
points(x,dpareto(x,xi=1.6), type="l", lty=3)  #now we are changing the $\xi$ so the tail is going steaper and steaper. 
points(x,dpareto(x,xi=2.5), type="l", lty=5)
abline(h=0)
abline(v=0)
```
\section{Generalized Extreme Values}

Changing the sequence
```{r}
x <- seq(from=-2.2, to=8, length=5000)
```

Ploting it \textbf{Generalized Extreme Values}

```{r}
plot(x, dgev(x,xi=.2), type="l", ylab="GEV densities", ylim=c(-.05,0.9))
points(x,dgev(x,xi=0.6), type="l", lty=3)  #now we are changing the $\xi$ so the tail is going steaper and steaper. 
points(x,dgev(x,xi=1.5), type="l", lty=5)
abline(h=0)
abline(v=0)
```

\section{Block Maxima}

```{r}
XX <- rpareto(1000000, xi=.4)
dim(XX) <- c(1000,1000)

MAX <- apply(XX, 2, max) #apply the max function on the matrix's collumn, you get a vector of max from each column. 2 - menas columns. 

qqplot(MAX, rgev(1000, xi=0.4)) #plots the MAX against the random of GEV, and MAX is heavier than... 
abline(0,1)
```

\section{Estimating the parameters}

MLE or MOM, if we have fat tail distribution look at Calculating L moments. p.83. 

Formula for L moments p.83

$\lambda_1= \alpha_0 = \int_0^1 F^{-1}(p)d $
$\lambda_2 = 2\alpha_1-\alpha_0 ... $

we are interested in the parameters and to verify if they are comparable to the theoretiacal ones mentioned in the book. 
If $m=0, \lambda=1, \xi=0$ then we are looking at the standard exponential distribution. 

```{r}
x <- rpareto(1000)
sample.LMOM(x)
```

In case we don't have enough observations:
```{r}
x <- rpareto(50, xi=0.4)

PPML <- plotting.positions(x)
PPML

SLM <- sample.LMOM(x)
SLM
```

L moments of GEV distribution

```{r}
x <- rgev(500, lambda=3.5, xi=0.4)
LMOMX <- sample.LMOM(x)
LMOMX
```

```{r}
x = rpareto(500, xi=0.4)
SLM <- sample.LMOM(x)
# gdp.lmom(SLM, location=0)
```

\section{Max likelihood estimation}

How to estimate these parameters. 

\textbf{For GEV}
```{r}
x <- rgev(500, lambda = 3.5, xi=0.4)
gev.ml(x)
```

\textbf{For GDP}
```{r}
y <- rpareto(500, lambda = 3.5, xi=0.4)
gpd.ml(y)
```


\section{Example PCS}

```{r}
data(PCS)
head(PCS)
PCS.index <- PCS[,2]
```

Fitting a GPD to the PCS.index data

```{r}
PCS.lmom <- gpd.lmom(PCS.index) 
```

Visualize
density = 20 makes the lines over the bars. 

```{r}
hist(PCS.index, breaks = 25, density = 20, freq = F)
```

Fix this code

```{r}
hist(PCS.index,breaks=25,density=20,freq=F) 
x <- seq(from=-1,to=160,length=1000)
points(x,dpareto(x,m=PCS.lmom$param.est[1],lambda=PCS.lmom$param.est[2],
                 xi=PCS.lmom$param.est[3]),type="l")
```

1. Check the goodness of fit in the tail 
2. generate large random sample from distribution fitted to the data 
3. produce qqq plot of Monte Carlo sample against original data set PCS. 

re: the chart - the tails are not well fitted. -> Method of Moments not as good as MAX likelihood. 

```{r}
PCS.ml <- gpd.ml(PCS.index)
PCS.ml <- PCS.ml$param.est
PCS.ml
```
As before, we are doing the monte Carlo

```{r}
PCS.rml <- rpareto(n=10000, m= PCS.lmom$param.est[1],
                   lambda = PCS.lmom$param.est[2],
                   xi = PCS.lmom$param.est[3])

qqplot(PCS.index, PCS.rml)
abline(0,1, col="red")
```
We are getting a good fit, they vary according to random sample - they are not fixed. 




Notes: 2.3 is being skipped

Q: is there a difference in estimating the VAR for portfolio?